---
title: "EDAV Final Project"
subtitle: |
  NYC Restaurant Inspection
  
author: "Cindy Wu, Selina Tang, Lisa Kim, Julie Yu"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    theme: paper
    toc: true
---
<style type="text/css">
body {
text-align: justify;
font-size: 12pt;
max-width: 1200px;
margin-left: 100px;
margin-right: 100px;
}
body .main-container {
max-width: 1200px;
font-size: 12pt;
}
</style>
## Introduction
(Explain why you chose this topic, and the questions you are interested in studying.
List team members and a description of how each contributed to the project.)

* In this project,...
* Team Members:


```{r echo=TRUE, message=FALSE}
library(knitr)
```

```{r packages, message=FALSE, warning=FALSE}
### Packages and Libraries
#<<packages_and_libraries>>
```


## Dataset: NYC Restaurant Inspection Data
* Our dataset is collected from NYC Open Data and can be directly accessed through - https://data.cityofnewyork.us/Health/DOHMH-New-York-City-Restaurant-Inspection-Results/xx67-kt59

* 375,989 rows of data entries of NYC restaurant information, violations and ratings collected between June 2012 - Mar 2018.

* Each row consists of (...)

## Data Quality
(Provide a detailed, well-organized description of data quality, including textual description, graphs, and code.)
```{r,eval = TRUE, fig.width=12, fig.height=4, fig.align='center', message=FALSE, warning=FALSE}
#code here
#sparsity, distribution,etc. 
```

## Exploratory Data Analysis
(Provide a detailed, well-organized description of your findings, including textual description, graphs, and code.  Your focus should be on both the results and the process. Include, as reasonable and relevant, approaches that didn't work, challenges, the data cleaning process, etc.)
```{r,eval = TRUE, fig.width=12, fig.height=4, fig.align='center', message=FALSE, warning=FALSE}
#code here
"
Scatter Plot
Histogram
Bar & Stack Bar Chart
Box Plot
Area Chart
Heat Map
Correlogram
"
```

## Executive Summary
(Provide a short nontechnical summary of the most revealing findings of your analysis  written for a nontechnical audience. The length should be approximately two pages (if we were using pages...) Take extra care to clean up your graphs, ensuring that best practices for presentation are followed.)

## Interactive Component 
```{r,eval = TRUE, fig.width=12, fig.height=4, fig.align='center', message=FALSE, warning=FALSE}
#code here
```

## Conclusion
